Learning features and abstract actions for computing generalized plans

### 2018-09-05 doc/paper.tex
Added Blai's changes

\newcommand{\hobs}{\ensuremath{h_{obs}}\xspace}
\newcommand{\indicator}[1]{[\![#1]\!]}
\newcommand{\C}{\mathcal{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\dexp}{\text{2-}\!\exp}

@@

\begin{abstract}
  Generalized planning is concerned with the   computation of  plans that solve multiple instances.
  Recently, it has been shown that such plans can be expressed as mappings of feature values into actions, and that they
  can often be  computed by means   of fully observable non-deterministic (FOND) planners after a sequence of
  transformations. The actions in such  plans, however,   are not the actions in the  instances themselves,  which  are
  not necessarily common to  other instances,   but  abstract actions that track the effect of the  actions
  on the given set of features. The formulation assumes  that both the features and the abstract actions are given.
  In this work, we address this limitation by  showing  how to learn them.  The resulting account of generalized
  planning  combines learning and planning in a novel way: a \emph{learner},  based on a min SAT formulation, yields
  the features and abstract actions  from    state transitions sampled  from some problem instances, 
  and a FOND \emph{planner} uses this information,   suitably   transformed as before,  to produce the general plans.
  Correctness guarantees are given and 
  experimental results on several  domains are reported.
\end{abstract}

@@

In the basic formulation due  to \citeay{hu:generalized}, a generalized plan
is a mapping of observations into actions that are assumed to be common
among  all the instances. More recently, this formulation has been extended by
\citeay{bonet:ijcai2018} to account for relational domains like Blocks world  where the set of
objects and  actions change from instance to instance.
For this,   the observations  are replaced by  a given  set of boolean
and numerical features $F$ and a set of \emph{abstract actions} $A_F$.
The  abstract actions are  \emph{sound} and \emph{complete} if they track the
effects  of the actions on  the features in a suitable way.
The resulting generalized plans  map feature values into abstract actions, and soundness
ensures that the application  of an   abstract action  results in the application of
a concrete action with a similar  effect  on  the features. Moreover, the form of the abstract actions
ensures that such plans can be computed using fully observable non-deterministic (FOND) planners
after some transformations.

@@

Bonet and Geffner's  formulation of generalized planning, however, assumes that the
features and the abstract actions are given.   In this work, we address this limitation.
We achieving this by  \emph{showing  how features and abstract actions  can be learned from
the primitive predicates used to define the instances and state transitions
sampled from them.} For example,  in \cite{bonet:ijcai2018}, 
the general policy for achieving the goal $clear(x)$  is obtained 
using a FOND planner by introducing the boolean feature $H$ for holding a block,
the numerical feature $n(x)$ for counting the  number of blocks above $x$, and  two   abstract  actions:
one  with  preconditions  $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement $n(x)$), the other with precondition $H$ and effect $\neg H$.
In this work, we show how  to obtain such  policies without
having to provide the  features and  abstract actions by hand.
The abstraction is learned instead from the domain predicates
and  state transitions sampled from some  instances.

The work has relation to a number of research threads in planning, knowledge representation
and machine learning.  We make use of SAT solvers and description logics for learning the features
and abstract actions \cite{sat-handbook,dl-handbook}. The abstract actions provide a model from which the plans are obtained via
transformations and FOND planners \cite{geffner:book,ghallab:book}.
The model is not the model of an instance but a generalized model for  obtaining  plans
that work for multiple instances. In this sense, the work is different than
action model learning \cite{yang:action-learning} and model-based reinforcement learning
and closer in aims to work on learning general policies from examples or experience
\cite{martin:concept,fern:bias,mazebase,general-drl}. Generalized planning has  also been formulated as a problem in first-order logic
\cite{srivastava:generalized} and general plans over finite horizons have also been derived using  first-order regression
\cite{boutilier2001symbolic,wang2008first,van2012solving}. Our approach differs from first-order approaches in the use
efficient propositional planners, and from purely learning approaches, in the \emph{formal guarantees} that are characteristic of planning:
if the learned abstract actions are actually sound,  the resulting general plan must be correct.

The paper is organized as follows. We provide the background first and  show then  how features and
abstract actions can be  learned by enforcing soundness and completeness over a set of samples  using 
a  pool of features obtained the   domain  predicates. The full  computational model  is then
summarized,  following by the experimental results and a discussion.

@@

A \emph{generalize planning problem} $\Q$ is a collection of planning instances $P$.
An instance $P$ is a  classical planning problems expressed in some compact language as a
tuple $P=\tup{V,I,A,G}$ where $V$ is a set of state variables that can take a finite set of values (boolean or not), $I$
is a set of atoms over $V$ defining an initial state $s_0$, $G$ is a set of atoms or literals over $V$ describing the goal states,
and $A$ is a set of actions $a$ with their preconditions and effects that define the set $A(s)$ of actions applicable
in any state $s$, and the successor state function $f(a,s)$, $a \in A(s)$. A state is a valuation over $V$, and a solution to
$P$ is an applicable action sequence $\pi=a_0,\ldots,a_n$ that generates  a state sequence $s_0,s_1,\ldots,s_{n}$
where $s_n$ is a goal state (makes $G$ true). In this sequence, $a_i \in A(s_i)$ and $s_{i+1}=f(a_i,s_i)$ for $i=0, \ldots, n-1$.
A state $s$ is reachable in $P$ if $s=s_n$ for one such sequence. A solution to the \emph{generalized problem} $\Q$ is
a solution to all instances $P \in Q$. The form of such solutions is described  below.

A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, 
and results in a value denoted as $\phi_f(s)$. 
A feature is \emph{boolean}  if it results in a boolean value, and  \emph{numeric} if it results
in a numerical value, that we assume to be integer and non-negative. The features $H$
and $n(x)$ represent two features in the Blocks world (BW): the first a boolean
feature encoding that   the gripper is not empty; the second, a numerical
feature encoding  the number of blocks above  $x$.

Symbols like $x$ to denote \emph{parameters} whose value depends on the instance  $P$ in  $\Q$.
For example, if $\Q_{clear}$  denotes  the BW instances with goals of the form $clear(x)$,
then the value of $x$ in an instance $P$ with goal $clear(a)$ is  $a$.

@@

Features $f$ refer to \emph{state functions} $\phi_f(s)$  in the instances of the generalized problem $\Q$,
but to \emph{state variables} in the abstraction.  Likewise, an \emph{abstract state} is  a truth valuation over the
state variables  $p$ associated with boolean features and the atoms $n=0$ over  state variables  $n$ associated with numerical features.
The abstract state $\bar{s}$  that \emph{corresponds} to a concrete state $s$ is the truth valuation that makes
$p$ true iff $\phi_p(s)=true$, and $n=0$ true iff $\phi_n(s) = 0$, and an abstract action $\bar{a}$
is applicable in a state $s$ if its preconditions are satisfied in $\bar{s}$.

A set of abstract actions $A_F$ is \emph{sound}  relative to a generalized problem $\Q$ 
iff for any reachable state $s$ over an instance $P$ in $\Q$, if an abstract action $\bar{a}$ in $A_F$
is applicable in $\bar{s}$, then there is a concrete action $a$ in $P$ that is applicable in $s$
that has the \emph{same qualitative  effects as}  $\bar{a}$ over the features. The action $a$
\emph{represents} or \emph{instantiates}  the abstract action $\bar{a}$ in $s$.

Similarly, a set of abstract actions $A_F$ is \emph{complete}
iff for any reachable state $s$ over an instance $P$ in $\Q$, if a concrete action $a$ in $P$
that is applicable in $s$, then there is an abstract action $\bar{a}$ in $A_F$
that is  applicable in $\bar{s}$ that has the \emph{same qualitative effects over the features as $a$.}

The actions $a$ and $\bar{a}$ \emph{have the same qualitative  effects} in a state $s$ 
when  they have the same   effects on the boolean features $p$,  and the same \emph{qualitative effects} on the numerical
features $n$; i.e., $a$ decreases (resp. increases) the value of the function $\phi_n(s)$ in $s$ iff
$\bar{a}$ decreases (resp. increases) the value of the state variable  $n$.

\begin{example}
  For the features $F=\{H,n(x)\}$, the abstract action  $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ 
  is sound in  problem $\Q_{clear}$ that includes the BW instances with stack and unstack actions.
  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is applicable in $\bar{s}$, 
  the gripper must be empty in $s$ and some block  must be above $x$. If $A$ is the clear block above $x$ and $on(A,B)$ is true in $s$,
  then the  action $a=unstack(A,B)$  will be  applicable in $s$ and have the same qualitative effects of  $\bar{a}$
  over $F$: they both make $H$ true and decrease the value of $n(x)$. The abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound.
\end{example}

For a set of abstract actions to be complete in this example, it should 
should contain  abstracts that track the  effects of the other  actions
on the features such as stacking a block above $x$, picking up a block that is not above $x$, and so on.
It may well be possible that a particular feature set $F$ can't support a set of abstract actions
$A_F$ that is both sound and complete relative to  $\Q$. Soundness, however,  is crucial for deriving general plans as it ensures
that in any state of any instance of $\Q$,  an applicable  abstract action  can be ``instantiated'' 
by   one or more concrete actions that have the same qualitative effects on the features.

@@

A solution to a  generalized problem $\Q$ given the  features $F$ and  abstract actions $A_F$
is a partial  function $\pi$ that maps \emph{abstract states} into \emph{abstract actions} such that
$\pi$ solves all instances $P$ in $\Q$. The plan or policy $\pi$ induces a  trajectory $s_0,a_0,s_1, \ldots, s_n$ in an instance $P$ of $\Q$
iff a)~$s_0$ is the initial state in $P$, b)~$a_i$ is one of the actions that instantiate 
$\pi(\bar{s_i})$ in $s_i$, c)~$s_{i+1}=f(a_i,s_i)$ if $a_i$ is applicable in $s_i$,
and d)~$s_n$ is a goal  state of $P$,  $n=\infty$, or $\pi(\bar{s_n})$ is undefined, denotes an abstract action
that is not applicable in $\bar{s_n}$, or no applicable action $a_n$ instantiates  $\pi(\bar{s_n})$.
The policy $\pi$ solves $P$ iff all the induced trajectories reach the goal of $P$.

\begin{example}
A solution for the generalized problem $\Q_{clear}$  with feature set $F=\{H,n(x)\}$
is the policy $\pi$  given by the rules $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
for the abstract actions $\bar{a}$ and $\bar{a'}$ above. The policy picks blocks above $x$ and puts them aside
(not above $x$) until $n(x)$ becomes zero.
\end{example}

@@

The steps for obtaining general policies $\pi$ for a generalized problem $\Q$
using the features $F$ and the  abstract actions $A_F$ are as follows \cite{bonet:ijcai2018}:

\medskip

\noindent 1. The state variables $V_F$ and the abstract actions $A_F$ are extended
  with initial and goal  formulas $I_F$ and $G_F$  over $V_F$ to yield a   \emph{numerical, non-deterministic planning problem}
  $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness,  $I_F$ must be  such that the initial states of instances $P$ in $\Q$ all satisfy $I_F$,
  while all states that satisfy $G_F$ must be goal states of $P$.\footnote{
   A state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}

\medskip

\noindent 2. The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a \emph{boolean FOND}
  problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by replacing 
  the numerical variables $n\in N$ by the symbols $n=0$, the first-order literals $n=0$ by propositional literals  $n=0$, 
  the effects $n\pplus$ by deterministic effects $n > 0$, and the effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.

  \medskip

\noindent 3. The strong cyclic solutions of  $Q'_F$ assume that the
  non-deterministic effects $n>0\,|\,n=0$ are \emph{fair} \cite{strong-cyclic},  
  yet  they are  \emph{conditionally fair} only: infinite decrements of $n$ ensure $n=0$  eventually but just without
  infinite  increments. The problem of obtanining  solutions of $Q'_F$ that assume conditional fairness only, the  terminating
  solutions in   \cite{srivastava:aaai2011}, is mapped into the problem of solving an  amended FOND $Q^+_F$
  \cite{bonet:ijcai2017}.\footnote{For this, we use a translation made available by the authors that fixes bugs in the  description found in the paper
     at https://github.com/bonetblai/qnp2fond.}

\medskip

The strong cyclic solutions of  $Q^+_F$ computed by FOND planners off-the-shelf   are solutions to the generalized problem $\Q_F$.

\begin{example}
  Let restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
  and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a'}\}$ as above, 
   $Q_F=\tup{V_F,I_F,G_F,A_F}$ can be defined with $I_F=\{\neg H, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
   $Q'_F$ is like $Q_F$ but with $n(x)=0$ regarded as a propositional symbol, $n(x) > 0$
  as its negation, and the effect $n(x)\mminus$ of  $\bar{a}$ replaced by  $\neg n(x)=0 \, | \, n(x)=0$.
  Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
  that corresponds to the policy above, solves $\Q_{clear}$.
\end{example}

@@

\begin{definition}
 For a generalized problem $\Q$, a \emph{sample set}  $\cal S$ is a non-empty set of tuples  $(s,s',P)$
 such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
 is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
 4)~$\cal S$ is closed in the sense  that if $(s,s',P)$ is in $\cal S$, then $(s,s'',P)$ is in $\cal S$  if $s''$ is another possible succesor of $s$ in $P$.
\end{definition}

By assuming that sampled states $s$ are tagged with the instance $P$, 
we can abbreviate the tuples $(s,s',P)$ as $(s,s')$, and refer to
the states $s'$ in  the pairs $(s,s') \in {\cal S}$ as the successors
of $s$ in ${\cal S}$. Notice that the closure condition 4) ensures that
states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in ${\cal S}$;
namely, all possible transitions $(s,s'')$ must be in the sample as well. We call them
the \emph{expanded} states in ${\cal S}$.

@@

\begin{definition}
  A set of abstract actions $A_F$ is sound \emph{relative to a  sample set} $\cal S$ for $\Q$
  iff for  any  abstract action $\bar{a}$ in $A_F$ that is applicable in  a expanded state $s$ of ${\cal S}$,
  there is a transition $(s,s')$ in $\cal S$ that has the same qualitative  effects over $F$ as $\bar{a}$.
\end{definition}

\begin{definition}
A set of abstract actions $A_F$ is complete   \emph{relative to a sample set} $\cal S$ for $\Q$
iff for each transition  $(s,s')$ in  $\cal S$, there is an abstract action $\bar{a}$ in $A_F$
with the same qualitative effects over in $F$ that is applicable in $s$.
\end{definition}

@@

For computing the features and abstract actions
we define  a  formula  $T({\cal S,F})$ that  is \emph{satisfiable} 
iff there is a set of features $F  \subseteq {\cal F}$ and a set of abstract actions $A_F$ over $F$
that is sound and complete relative to the sample set ${\cal S}$.

\Omit{The  sets  $F$ and $A_F$ can
     then be read off  from any  satisfying assignment. Later we will add weights
     to so
     me of the clauses to establish a preference over these assignments
     (e.g., for  minimizing $|F|$).}

@@

For each transition  $(s, s') \in \mathcal{S}$ and each feature $f \in \mathcal{F}$,
$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the \emph{qualitative change of value}
of feature $f$ along the transition $(s, s')$, which can go from false to true ($+$), from true to false ($-$) or remain unchanged ($\bot$),
for boolean features, and can increase ($\uparrow$), decrease ($\downarrow$), or remain unchanged ($\bot$), for numerical features.
The \emph{propositional variables}  in $T({\cal S,F})$ are then:

>>>>

\newcommand{\hobs}{\ensuremath{h_{obs}}\xspace}
\newcommand{\indicator}[1]{[\![#1]\!]}
\newcommand{\C}{\mathcal{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\F}{\mathcal{F}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\dexp}{\text{2-}\!\exp}

@@

\begin{abstract}
  Generalized planning is concerned with the computation of plans that solve multiple instances.
  Recently, it has been shown that generalized plans can be expressed as mappings of feature
  values into actions, and that they can often be computed with fully observable non-deterministic
  (FOND) planners on suitable FOND problems. %after a sequence of transformations.
  The actions in such plans, however, are not the actions in the instances themselves, which are
  not necessarily common to  other instances, but abstract actions that track the effect of the
  actions on the set of features. The formulation assumes that both the features and the abstract
  actions are given.
  In this work, we address this limitation by showing how to automatically learn the features and
  abstract actions. The resulting account of generalized planning combines learning and planning
  in a novel way: a \emph{learner}, based on a min SAT formulation, yields the features and abstract
  actions from state transitions sampled from some problem instances, and a FOND \emph{planner}
  uses this information, suitably transformed, to produce the general plans.
  Correctness guarantees are given and experimental results on several domains are reported.
\end{abstract}

@@

In the basic formulation due  to \citeay{hu:generalized}, a generalized plan
is a mapping of observations into actions that are assumed to be common
among  all the instances. More recently, this formulation has been extended by
\citeay{bonet:ijcai2018} to account for relational domains like Blocksworld  where the set of
objects and  actions change from instance to instance.
In the new formulation, the observations  are replaced by  a set of boolean
and numerical features $F$ and a set of \emph{abstract actions} $A_F$.
The  abstract actions are  \emph{sound} and \emph{complete} if they track the
effects  of the actions on  the features in a suitable way.
The resulting generalized plans  map feature values into abstract actions, and soundness
ensures that the application  of an   abstract action  results in the application of
a concrete action with a similar  effect  on  the features. Moreover, the form of the abstract actions
ensures that generalized plans can be computed using fully observable non-deterministic (FOND) planners
once the generalized planning problem is transformed into a FOND problem.
%after some transformations.

@@

Bonet and Geffner's  formulation of generalized planning assumes that the
features and abstract actions are given. In this work we address this limitation
by  \emph{showing  how the features and abstract actions  can be learned from
the primitive predicates used to define the instances and a given sample of state transitions.}
For example,  the general policy for achieving $clear(x)$ of \citeay{bonet:ijcai2018}
is obtained using a FOND planner on an abstracton that consists of the boolean
feature $H$ that tracks whether the gripper holds a block and a numerical feature
$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
In this work, we show how to obtain such policies without having to provide the
features and abstract actions by hand.  The abstraction is learned instead from
the domain predicates and state transitions sampled from some instances.

The work has relation to a number of research threads in planning, knowledge
representation, and machine learning. We make use of SAT solvers and description
logics for learning features and abstract actions \cite{sat-handbook,dl-handbook}.
The abstract actions provide a model from which the plans are obtained via
transformations and FOND planners \cite{geffner:book,ghallab:book}.
The model is not the model of an instance but a generalized model for obtaining
plans that work for multiple instances. In this sense, the work is different than
action model learning \cite{yang:action-learning} and model-based reinforcement
learning and closer in aims to work on learning general policies from examples or
experience \cite{martin:concept,fern:bias,mazebase,general-drl}.
Generalized planning has also been formulated as a problem in first-order logic
\cite{srivastava:generalized} and general plans over finite horizons have been
derived using first-order regression \cite{boutilier2001symbolic,wang2008first,van2012solving}.
Our approach differs from first-order approaches in the use of propositional
planners, and from purely learning approaches, in the \emph{formal guarantees} that
are characteristic of planning: if the learned abstract actions are actually sound,
the resulting general plans are correct.

The paper is organized as follows. We provide the background first and show
how features and abstract actions can be learned by enforcing soundness and
completeness over a set of samples using a pool of features obtained the
domain predicates. The full computational model is then summarized, followed
by experimental results and a discussion.

@@

A \emph{generalized planning problem} $\Q$ is a collection of planning instances $P$.
An instance $P$ is a  classical planning problems expressed in some compact language
as a tuple $P=\tup{V,I,G,A}$ where $V$ is a set of state variables that can take a
finite set of values (boolean or not), $I$ is a set of atoms over $V$ defining an
initial state $s_0$, $G$ is a set of atoms or literals over $V$ describing the goal
states, and $A$ is a set of actions $a$ with their preconditions and effects that
define the set $A(s)$ of actions applicable in any state $s$, and the successor
state function $f(a,s)$, for any state $s$ and $a \in A(s)$.
A state is a valuation over $V$, and a solution to $P$ is an applicable action
sequence $\pi=a_0,\ldots,a_n$ that generates  a state sequence $s_0,s_1,\ldots,s_{n}$
where $s_n$ is a goal state (makes $G$ true). In this sequence, $a_i \in A(s_i)$
and $s_{i+1}=f(a_i,s_i)$ for $i=0, \ldots, n-1$. A state $s$ is reachable in $P$
if $s=s_n$ for one such sequence. A solution to the \emph{generalized problem}
$\Q$ is a solution to all instances $P \in Q$. The form of such solutions is
described below.

A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
in a value denoted as $\phi_f(s)$. 
A feature is \emph{boolean} if it results in a boolean value, and \emph{numeric}
if it results in a numerical value, that we assume to be a non-negative integer.
The features $H$ and $n(x)$ represent two features in the Blocksworld:
the first is a boolean feature that tracks whether the gripper is empty, while
the second is numerical and tracks the number of blocks above $x$.

Symbols like $x$ denote \emph{parameters} whose value depends on the
instance $P$ in $\Q$. For example, if $\Q_{clear}$ denotes the Blocksworld
instances with goals of the form $clear(x)$, the value of $x$ in an instance
$P$ with goal $clear(a)$ is $a$.

@@

Features $f$ refer to \emph{state functions} $\phi_f(s)$ in the instances
of the generalized problem $\Q$, but to \emph{state variables} in the
abstraction. Likewise, an \emph{abstract state} is  a truth valuation over
the state variables $p$ associated with the boolean features and the atoms
$n=0$ over the state variables $n$ associated with the numerical features.
The abstract state $\bar{s}$ that \emph{corresponds} to a concrete state $s$
is the truth valuation that makes $p$ true iff $\phi_p(s)=true$, and $n=0$
true iff $\phi_n(s) = 0$, and an abstract action $\bar{a}$ is applicable in
a state $s$ if its preconditions are satisfied in $\bar{s}$.

A set of abstract actions $A_F$ is \emph{sound} relative to a generalized
problem $\Q$ iff for any reachable state $s$ over an instance $P$ in $\Q$
and any abstract action $\bar{a}$ in $A_F$ that is applicable in $\bar{s}$,
there is a concrete action $a$ in $P$ that is applicable in $s$ and that
has the \emph{same qualitative effects as} $\bar{a}$ over the features.
The action $a$ \emph{represents} or \emph{instantiates} the abstract action
$\bar{a}$ on $s$. 
%
Similarly, a set of abstract actions $A_F$ is \emph{complete} iff for any
reachable state $s$ over an instance $P$ in $\Q$ and any concrete action $a$
in $P$ that is applicable in $s$, there is an abstract action $\bar{a}$ in
$A_F$ that is applicable in $\bar{s}$ and that has the \emph{same qualitative
effects over the features as $a$.}

The actions $a$ and $\bar{a}$ \emph{have the same qualitative effects} in
a state $s$ when they have the same effects on the boolean features $p$,
and the same \emph{qualitative effects} on the numerical features $n$; i.e.,
$a$ decreases (resp.\ increases) the value of the function $\phi_n(s)$ in
$s$ iff $\bar{a}$ decreases (resp.\ increases) the value of the state
variable  $n$.

\begin{example}
  For the features $F=\{H,n(x)\}$, the abstract action $\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$ 
  is sound in the generalized problem $\Q_{clear}$ which consists of all
  the Blocksworld instances, with stack and unstack actions, whose goal
  is $clear(x)$.
  Indeed, if $s$ is a reachable state in an instance $P$ and $\bar{a}$ is
  applicable in $\bar{s}$, the gripper must be empty in $s$ and some
  block must be above $x$. Thus, if $A$ is the clear block above $x$ and
  $on(A,B)$ is true in $s$, the action $a=unstack(A,B)$ is applicable in
  $s$ and it has the same qualitative effects of $\bar{a}$ over $F$: they
  both make $H$ true and both decrease the value of $n(x)$.
  Likewise, the abstract action $\bar{a'}=\abst{H}{\neg H}$ is also sound. \qed
\end{example}

For a set of abstract actions to be complete in this example, it should 
contain abstracts actions to represent all the concrete actions in the
problem such as stacking a block above $x$, picking up a block that is
not above $x$, and so on.
It may well be possible that a particular feature set $F$ cannot support
a set of abstract actions $A_F$ that is both sound and complete relative
to $\Q$. Soundness, however, is crucial for deriving general plans as it
ensures that in any state of any instance of $\Q$, an applicable abstract
action  can be ``instantiated'' by one or more concrete actions that have
the same qualitative effects on the features.

@@

A solution to a generalized problem $\Q$ given the features $F$ and abstract
actions $A_F$ is a partial function $\pi$ that maps \emph{abstract states}
into \emph{abstract actions} such that $\pi$ solves all instances $P$ in $\Q$.
The plan or policy $\pi$ induces a trajectory $s_0,a_0,s_1, \ldots, s_n$ in
an instance $P$ of $\Q$ iff
1)~$s_0$ is the initial state in $P$,
2)~$a_i$ is one of the actions that instantiate $\pi(\bar{s}_i)$ in $s_i$,
3)~$a_i$ is applicable in $s_i$ and $s_{i+1}=f(a_i,s_i)$, and
4)~$s_n$ is a goal state of $P$, $n=\infty$, or $\pi(\bar{s}_n)$ either is
undefined, denotes an abstract action that is not applicable in $\bar{s_n}$,
or no applicable action $a_n$ instantiates $\pi(\bar{s}_n)$.
The policy $\pi$ solves $P$ iff all the induced trajectories reach the
goal of $P$.

\begin{example}
  A solution for $\Q_{clear}$ with feature set
  $F=\{H,n(x)\}$ is the policy $\pi$ given by the rules
  $\Rule{\neg H, n(x)>0}{\bar{a}}$ and $\Rule{H, n(x)>0}{\bar{a}'}$
  for the abstract actions $\bar{a}$ and $\bar{a'}$ above.
  The policy picks blocks above $x$ and puts them aside (not above $x$)
  until $n(x)$ becomes zero. \qed
\end{example}

@@

The steps for obtaining policies $\pi$ for a generalized problem $\Q$
using the features $F$ and abstract actions $A_F$ are as follows \cite{bonet:ijcai2018}:

\begin{enumerate}[1.]
  \item The state variables $V_F$ and the abstract actions $A_F$ are
    extended with initial and goal formulas $I_F$ and $G_F$ over $V_F$
    to yield a \emph{numerical and non-deterministic planning problem}
    $Q_F=\tup{V_F,I_F,G_F,A_F}$. For soundness, $I_F$ must be such that
    the initial states of instances $P$ in $\Q$ all satisfy $I_F$, while
    all states that satisfy $G_F$ must be goal states of $P$.\footnote{A
    state $s$ is said to satisfy $I_F$ or $G_F$ if  $\bar{s}$ does.}
%
  \item The problem $Q_F=\tup{V_F,I_F,G_F,A_F}$ is converted into a
    \emph{boolean FOND} problem $Q'_F=\tup{V'_F,I'_F,G'_F,A'_F}$ by
    replacing the numerical variables $n\in N$ by the symbols $n=0$,
    the first-order literals $n=0$ by propositional literals  $n=0$, 
    the effects $n\pplus$ by deterministic effects $n > 0$, and the
    effects $n\mminus$ by non-deterministic effects $n>0\,|\,n=0$.
%
  \item The solutions computed by a FOND planner on $Q'_F$ are the
    strong cyclic solutions of $Q'_F$ \cite{strong-cyclic}.
    Such solutions however do not necessarily solve $Q_F$ because the
    non-determistic effects $n>0\,|\,n=0$ in $Q'_F$ are \emph{conditionally fair}
    instead of being just \emph{fair}: infinite decrements of $n$ ensure
    $n=0$ eventually \emph{only} when number of increments of $n$ is finite.
    The problem of obtanining solutions of $Q'_F$ that only assume conditional
    fairness, so-called terminating solutions in \cite{srivastava:aaai2011},
    is mapped into the problem of solving an amended FOND $Q^+_F$ that
    only assumes (standard) fairness \cite{bonet:ijcai2017}.\footnote{The
    translation to $Q^+_F$ is performed with software kindly supplied by the authors of \cite{bonet:ijcai2017}.}
    %available by the authors that fixes bugs in the description found
    %in the paper at https://github.com/bonetblai/qnp2fond.}
\end{enumerate}

%The strong cyclic solutions of  $Q^+_F$ computed by FOND planners off-the-shelf   are solutions to the generalized problem $\Q_F$.

\begin{example}
  Let restrict  $\Q_{clear}$  to those  instances where  the gripper is initially empty
  and there are blocks on top of $x$. For $F=\{H,n(x)\}$  and $A_F=\{\bar{a},\bar{a'}\}$ as above, 
  $Q_F=\tup{V_F,I_F,G_F,A_F}$ can be defined with $I_F=\{\neg H, n(x) > 0\}$ and $G_F=\{n(x)=0\}$.
  $Q'_F$ is like $Q_F$ but with $n(x)=0$ regarded as a propositional symbol, $n(x) > 0$
  as its negation, and the effect $n(x)\mminus$ of  $\bar{a}$ replaced by  $\neg n(x)=0 \, | \, n(x)=0$.
  Since  no action in $A_F$  increases $n(x)$,   $Q^+_F=Q'_F$, and the strong cyclic solution  of $Q'_F$,
  that corresponds to the policy above, solves $\Q_{clear}$. \qed
\end{example}

@@

\begin{definition}
 For a generalized problem $\Q$, a \emph{sample set}  $\S$ is a non-empty set of tuples  $(s,s',P)$
 such that 1)~$P$ is an instance of $\Q$, 2)~$s$ is a reachable state in $P$, 3)~$s'$
 is a possible succesor of $s$ in $P$; i.e., $s'=f(a,s)$ for some action $a$ applicable in $s$,
 4)~$\S$ is closed in the sense  that if $(s,s',P)$ is in $\S$, then $(s,s'',P)$ is in $\S$  if $s''$ is another possible succesor of $s$ in $P$.
\end{definition}

By assuming that sampled states $s$ are tagged with the instance $P$, 
we abbreviate the tuples $(s,s',P)$ as $(s,s')$, and refer to the
states $s'$ in  the pairs $(s,s') \in \S$ as the successors of
$s$ in $\S$. Notice that the closure condition ensures that
states $s$ appearing \emph{first} in pairs $(s,s')$ must be fully expanded in $\S$;
namely, all possible transitions $(s,s'')$ must be in the sample as well. We call them
the \emph{expanded} states in $\S$.

@@

\begin{definition}
  A set of abstract actions $A_F$ is sound \emph{relative to a  sample set} $\S$ for $\Q$
  iff for  any  abstract action $\bar{a}$ in $A_F$ that is applicable in  a expanded state $s$ of $\S$,
  there is a transition $(s,s')$ in $\S$ that has the same qualitative  effects over $F$ as $\bar{a}$.
\end{definition}

\begin{definition}
  A set of abstract actions $A_F$ is complete   \emph{relative to a sample set} $\S$ for $\Q$
  iff for each transition  $(s,s')$ in  $\S$, there is an abstract action $\bar{a}$ in $A_F$
  with the same qualitative effects over in $F$ that is applicable in $s$.
\end{definition}

@@

For computing the features and abstract actions we define a formula
$T(\S,\F)$ that is \emph{satisfiable} iff there is a set of
features $F \subseteq \F$ and a set of abstract actions $A_F$
over $F$ that is sound and complete relative to the sample set
$\S$.
\textcolor{red}{\bf **** Need to say what $\F$ is ****}

\Omit{The  sets  $F$ and $A_F$ can then be read off from any satisfying
  assignment. Later we will add weights to so me of the clauses to establish
  a preference over these assignments (e.g., for  minimizing $|F|$).
}

@@

For each transition  $(s, s') \in \S$ and each feature $f \in \F$,
$\Delta_f(s, s') \in \{+, -, \uparrow, \downarrow, \bot\}$ will denote the
\emph{qualitative change of value} of feature $f$ along the transition $(s, s')$,
which can go from false to true ($+$), from true to false ($-$) or remain
unchanged ($\bot$), for boolean features, and can increase ($\uparrow$),
decrease ($\downarrow$), or remain unchanged ($\bot$), for numerical features.
The \emph{propositional variables} in $T(\S,\F)$ are then:
\begin{enumerate}[{\small$\bullet$}]
  \item $selected(f)$ for each $f \in \F$, true iff $f$ selected (in $F$).
  \item $D_1(s,t)$ for states $s$ and $t$ expanded in $\S$,  true iff
    selected features distinguish $s$ from $t$; i.e.\ if $s$ and $t$ disagree
    on the truth value of a selected boolean feature $p$ or atom $n=0$ for a
    selected numerical feature $n$.
    %Since $D_1$ is symmetric, only symbols for $s < t$ are created for some
    %arbitrary, fixed order over the states in $\S_1$.
  \item $D_2(s, s', t, t')$ for each $(s, s')$ and $(t, t')$ in $\S$,
    true iff selected features $f$ distinguish the two transitions; i.e., if
    $\Delta_f(s, s')\not=\Delta_f(t,t')$. 
    %As for $D_1$, the number of $D_2$ variables is cut in half due to symmetries.
\end{enumerate}

\medskip
\noindent The first formulas in $T(\S,\F)$ capture the meaning of $D_1$:

\begin{equation}
  \label{eq:d1}
  D_1(s, t) \ \Leftrightarrow\ \textstyle \bigvee_{f}  selected(f)
\end{equation}

\noindent where $f$ ranges over features in $\F$ with  different qualitative
values in $s$ and $t$; i.e.\ boolean features $p$ with different truth value
in $s$ and $t$, and numerical features $n$ with different truth values for
atom $n=0$ in $s$ and $t$.
